---
title: "Data processing"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r include=FALSE}

knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE)

```

# Dependencies 

```{r message=FALSE, warning=FALSE}

library(tidyverse)

```

# Tidying

```{r}

# read in and wrangle data
data_tidied <- 
  bind_rows(read_csv(file = "../raw/Genova_ML5.csv") %>%
              mutate(site = "Genova",
                     country = "Italy",
                     unique_id = paste0(site, subject),
                     stimulusitem5 = as.character(stimulusitem5)),
            read_csv(file = "../raw/Milan_ML5.csv") %>%
              mutate(site = "Milan",
                     country = "Italy",
                     unique_id = paste0(site, subject),
                     stimulusitem5 = as.character(stimulusitem5)),
            read_csv(file = "../raw/Rome_ML5.csv") %>%
              mutate(site = "Rome",
                     country = "Italy",
                     unique_id = paste0(site, subject),
                     stimulusitem5 = as.character(stimulusitem5)),
            read_csv(file = "../raw/UCF_ML5.csv") %>%
              mutate(site = "UCF",
                     country = "USA",
                     unique_id = paste0(site, subject),
                     stimulusitem5 = as.character(stimulusitem5)),
            read_csv(file = "../raw/UVA_ML5.csv") %>%
              mutate(site = "UVA",
                     country = "USA",
                     unique_id = paste0(site, subject),
                     stimulusitem5 = as.character(stimulusitem5)),
            read_csv(file = "../raw/Padova_ML5.csv") %>%
              mutate(site = "Padova",
                     country = "Italy",
                     unique_id = paste0(site, subject),
                     stimulusitem5 = as.character(stimulusitem5)),
            read_csv(file = "../raw/VCU_ML5.csv") %>%
              mutate(site = "VCU",
                     country = "USA",
                     unique_id = paste0(site, subject),
                     stimulusitem5 = as.character(stimulusitem5)),
            read_csv(file = "../raw/PLU_ML5.csv") %>%
              mutate(site = "PLU",
                     country = "USA",
                     unique_id = paste0(site, subject),
                     stimulusitem5 = as.character(stimulusitem5))) %>%
  
  # add experiment condition - not in data, derived from subject 
  mutate(condition = ifelse(subject %% 2 == 0, "high social pressure",
                            ifelse(subject %% 2 == 1, "low social pressure", NA))) %>%
  
  # completeness
  group_by(unique_id) %>%
  mutate(n_rows = n(),
         complete_data = ifelse(n_rows == 100, TRUE, FALSE)) %>%
  ungroup() %>%
  
  # exclusions
  # because 34 and 162 because they could read the Chinese characters
  filter(!unique_id %in% c("UVA834", "UVA962")) %>%
  # subject < 3400 (5000-1600) ### in ML5 but not explained??
  filter(subject < 3400) %>%
  
  # recoding
  mutate(rating = case_when(response == "minus2" ~ 1,
                            response == "minus1" ~ 2,
                            response == "plus1" ~ 3,
                            response == "plus2" ~ 4)) %>%
  mutate(evaluative_task = ifelse(trialcode %in% c("whitedirect", "blackdirect"), "direct", 
                                  ifelse(trialcode %in% c("whiteindirect", "blackindirect"), "indirect", NA)),
         prime_type = ifelse(trialcode %in% c("whitedirect", "whiteindirect"), "white", 
                             ifelse(trialcode %in% c("blackdirect", "blackindirect"), "black", NA)))

```

# Validation checks

```{r}

data_tidied %>%
  distinct(unique_id, .keep_all = TRUE) %>%
  count(complete_data)

data_tidied %>%
  distinct(unique_id, .keep_all = TRUE) %>%
  count(n_rows) %>%
  arrange(desc(n))

```

- 100 participants have incomplete or excess data. 50 appear to have doubled data; possibly participant codes that got reused and are actually two participants. explore their demographics to see if they are duplicate same participants or different participants, i.e., can this data be saved somehow?

# Additional exclusions 

- compared to ML5, I added data exclusions for data completeness and for participant codes < 1. 

```{r}

data_after_exclusions <- data_tidied %>%
# missing participant code ### WASNT IN ML5
  filter(!is.na(subject)) %>%
  # subject is not a demo ### WASNT IN ML5
  filter(subject > 0) %>%
  # data completeness ### NOT INCLUDED IN ML5
  filter(complete_data == TRUE)

```

# Conditions

```{r}

data_conditions <- data_after_exclusions %>%
  distinct(unique_id, condition, site, country)

```

# Demographics

```{r}

data_demographics <- data_after_exclusions %>%
  # filter out only the demogrpahics items
  filter(blockcode == "demo") %>%
  select(unique_id, trialcode, response) %>%
  # recode italian to english
  mutate(trialcode = case_when(trialcode == "anni" ~ "age",
                               trialcode == "etnia" ~ "ethnicity",
                               trialcode == "sesso" ~ "gender",
                               TRUE ~ trialcode)) %>%
  # reshape to wide format
  pivot_wider(names_from = trialcode,
              values_from = response) %>%
  # recode from numeric to factor
  mutate(gender = case_when(gender == 1 ~ "male",  # codings validated against the inquisit script
                            gender == 2 ~ "female",
                            gender == 3 ~ "other",
                            gender == 4 ~ "prefer not to say")) %>%
  # recode age to english
  mutate(age = case_when(age == "pi?? di 45" ~ "over 45",
                         TRUE ~ age))

```

- NB rescoring of ethnicity from italian to english still needs to be done

# Scoring of direct and indirect evaluative measures

```{r}

data_evaluations <- data_after_exclusions %>%
  # filter out only the evaluations tasks
  filter(evaluative_task %in% c("direct", "indirect")) %>%
  # calculate mean ratings for each task and prime type
  group_by(unique_id, condition, evaluative_task, prime_type) %>%
  summarize(mean_rating = mean(rating), .groups = "drop") %>%
  # wrangle the data to wide format and clarify names
  pivot_wider(names_from = prime_type,
              values_from = mean_rating) %>%
  rename(mean_rating_white_primes = white,
         mean_rating_black_primes = black) %>%
  mutate(diff_in_mean_ratings = mean_rating_white_primes - mean_rating_black_primes) %>%
  pivot_wider(names_from = evaluative_task,
              values_from = c(diff_in_mean_ratings, 
                              mean_rating_white_primes,
                              mean_rating_black_primes))

```

# Combine and write to disk

```{r}

data_processed <- data_conditions %>%
  full_join(data_demographics, by = "unique_id") %>%
  full_join(data_evaluations, by = c("unique_id", "condition")) %>%
  select(unique_id, 
         country,                          
         site, 
         condition,                        
         gender, 
         age,                              
         ethnicity, 
         diff_in_mean_ratings_direct,      
         diff_in_mean_ratings_indirect, 
         mean_rating_white_primes_direct,  
         mean_rating_white_primes_indirect,
         mean_rating_black_primes_direct,  
         mean_rating_black_primes_indirect)

write_csv(data_processed, "data_processed.csv")

```

# Things included in ML5 code not included here:

- Ratio scoring. Instead of differences in means, Vianello et al. 2015 used ratio scores: mean_rating_white_primes / mean_rating_black_primes. Because this is uncommon in the AMP literature I don't do it here.

# Session info

```{r}

sessionInfo()

```


