---
title: "Analyses"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r include=FALSE}

knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE)

# turn off scientific notation
options(scipen = 999)

```

# Dependencies 

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(lme4)
library(lmerTest)
library(MuMIn)
library(sjPlot)
library(brms)
library(parallel)
library(bayestestR)

```

# Data

```{r}

data_processed <- read.csv("../data/processed/data_processed.csv") %>%
  mutate(country = as.factor(country),
         site = as.factor(site),
         condition = fct_relevel(as.factor(condition), "low social pressure", "high social pressure"),
         age = as.character(age), # NB age is a character as ">45" was the top category
         gender = as.factor(gender),
         ethnicity = as.factor(ethnicity)) %>%
  rename(indirect = diff_in_mean_ratings_indirect,
         direct = diff_in_mean_ratings_direct)

```

# Changes from ML5 analyses

ML5's models: 

- Did not exclude for missing or duplicate data. This was corrected: 5% of data was excluded on this basis.
- Stated that data collection at the Italian vs. US sites was considered meaningful (i.e., concerns raised by Payne about data collection in Italian motivated this replication to be run). Non-random assignment to Italian vs USA data collection was therefore important to acknowledge in the models' random structures, but was not in the ML5's models. This was corrected.
- Found singular model fits. Bayesian modeling via MCMC may allow for better model convergence.
- Did not allow a random slope for indirect (ie assumed implicit-explicit correlations were identical between sites). Inspection of the site level data suggests these correlations did indeed vary significantly (see below table). This was corrected.

```{r}

data_processed %>%
  group_by(site) %>%
  summarize(n = n(),
            implicit_explicit_correlation = round(cor(direct, indirect), 2)) %>%
  arrange(desc(implicit_explicit_correlation))

```

# Testing primary replication effect

## Fit models

```{r}

# Step 1 model - all but interaction
fit_1 <- lmer(direct ~ indirect + condition + (indirect + condition | country/site),
              REML = FALSE,
              data = data_processed)

# Step 2 model - adds interaction
fit_2 <- lmer(direct ~ indirect * condition + (indirect + condition | country/site),
              REML = FALSE,
              data = data_processed)

```

## Inspect model 2 (including interaction)

```{r}

tab_model(fit_2, 
          show.std = TRUE,
          emph.p = FALSE,
          col.order = c("est", "ci", "std.est", "std.ci", "p"))

```

## ANOVA

```{r}

anova(fit_1, fit_2)

```

## $R^2$ for interaction effect

```{r}

r.squaredLR(fit_2, fit_1)

```

## Plot predictions

```{r fig.height=6, fig.width=7.5, message=FALSE, warning=FALSE}

plot_model(fit_2, type = "int") +  # , show.data = TRUE
  xlim(-2, 2) +
  ylim(-2, 2) +
  scale_color_viridis_d(begin = 0.3, end = 0.7) +
  scale_fill_viridis_d(begin = 0.3, end = 0.7) +
  xlab("Indirect measure\n(Instruction to evaluate primes)") +
  ylab("Direct measure\n(Instruction to evaluate targets)") +
  ggtitle("")

```

# Moderation by country

Payne argued that conducting the replication in Italy and in Italian was problematic. 

```{r}

fit_3 <- lmer(direct ~ indirect * condition * country + (indirect + condition | site),
              REML = FALSE,
              data = data_processed)

tab_model(fit_3, 
          emph.p = FALSE,
          col.order = c("est", "ci", "std.est", "std.ci", "p"))

plot_model(fit_3, type = "int") +  # , show.data = TRUE
  xlim(-2, 2) +
  ylim(-2, 2) +
  scale_color_viridis_d(begin = 0.3, end = 0.7) +
  scale_fill_viridis_d(begin = 0.3, end = 0.7) +
  xlab("Indirect measure\n(Instruction to evaluate primes)") +
  ylab("Direct measure\n(Instruction to evaluate targets)") +
  ggtitle("")

```


# Bayesian models

Given singular fit, comparable Bayesian models were fit as a sensitivity analysis.  

## Fit models

```{r}

# create needed directory
dir.create("models")

# Step 1 model - all but interaction
fit_1_bayesian <- brm(direct ~ indirect + condition + (indirect + condition | country/site),
                      data      = data_processed,
                      file      = "models/fit_1_bayesian",
                      prior     = prior(normal(0, 2)),
                      iter      = 13000,
                      warmup    = 3000,
                      control   = list(adapt_delta = 0.95),  # to avoid divergent transitions
                      chains    = 4,
                      cores     = parallel::detectCores(),
                      save_pars = save_pars(all = TRUE))

# Step 2 model - adds interaction
fit_2_bayesian <- brm(direct ~ indirect * condition + (indirect + condition | country/site),
                      data      = data_processed,
                      file      = "models/fit_2_bayesian",
                      prior     = prior(normal(0, 2)),
                      iter      = 13000,
                      warmup    = 3000,
                      control   = list(adapt_delta = 0.95),  # to avoid divergent transitions
                      chains    = 4,
                      cores     = parallel::detectCores(),
                      save_pars = save_pars(all = TRUE))

add_criterion(fit_1_bayesian, criterion = c("loo", "bayes_R2"))
add_criterion(fit_2_bayesian, criterion = c("loo", "bayes_R2"))

```

## Inspect model 2 (including interaction)

```{r}

tab_model(fit_2_bayesian, 
          show.std = TRUE,
          emph.p = FALSE)

```

## LOO

```{r}

loo(fit_1_bayesian)
loo(fit_2_bayesian)

```

## Bayes factor

```{r}

summary(fit_2_bayesian)

bayesfactor_models(fit_2_bayesian, denominator = fit_1_bayesian)

```

## $R^2$ for interaction effect

```{r}

bind_rows(as_tibble(bayes_R2(fit_1_bayesian)) %>% mutate(model = "fit_1"),
          as_tibble(bayes_R2(fit_2_bayesian)) %>% mutate(model = "fit_2")) %>%
  select(model, Beta = Estimate, CI_lower = Q2.5, CI_upper = Q97.5) %>%
  mutate_if(is.numeric, round, digits = 3)

```

## Plot predictions

```{r fig.height=6, fig.width=7.5, message=FALSE, warning=FALSE}

# plot_model(fit_2_bayesian, type = "int") +  # , show.data = TRUE
#   xlim(-2, 2) +
#   ylim(-2, 2) +
#   scale_color_viridis_d(begin = 0.3, end = 0.7) +
#   xlab("Indirect measure\n(Instruction to evaluate primes)") +
#   ylab("Direct measure\n(Instruction to evaluate targets)") +
#   ggtitle("")

```

## Simplified models

```{r}

# Step 1 model - all but interaction
fit_3_bayesian <- brm(direct ~ indirect + condition,
                      data    = data_processed,
                      file    = "models/fit_3_bayesian",
                      prior   = prior(normal(0, 1)),
                      iter    = 6000,
                      warmup  = 3000,
                      control = list(adapt_delta = 0.95),  # to avoid divergent transitions
                      chains  = 4,
                      cores   = parallel::detectCores())

# Step 2 model - adds interaction
fit_4_bayesian <- brm(direct ~ indirect * condition,
                      data    = data_processed,
                      file    = "models/fit_4_bayesian",
                      prior   = prior(normal(0, 1)),
                      iter    = 6000,
                      warmup  = 3000,
                      control = list(adapt_delta = 0.95),  # to avoid divergent transitions
                      chains  = 4,
                      cores   = parallel::detectCores())

add_criterion(fit_3_bayesian, criterion = c("bayes_R2"))
add_criterion(fit_4_bayesian, criterion = c("bayes_R2"))

bind_rows(as_tibble(bayes_R2(fit_3_bayesian)) %>% mutate(model = "fit_3"),
          as_tibble(bayes_R2(fit_4_bayesian)) %>% mutate(model = "fit_4")) %>%
  select(model, Beta = Estimate, CI_lower = Q2.5, CI_upper = Q97.5) %>%
  mutate_if(is.numeric, round, digits = 3)

plot_model(fit_4_bayesian)

```

# Session info

```{r}

sessionInfo()

```



